---
title: "HW #2"
author: "Andrew Shao"
date: "2024-09-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### (a)
$n$ is the number of measurements taken, $Y_{n \times 1}$ is the vector of observed measurements, and $X_{n \times p}$ is the matrix of 1's and 0's corresponding to whether or not each ball was weighed for the measurement.
\[n = 5, 
Y_{n \times 1} = \begin{pmatrix} 2 \\ 1 \\ 4 \\ 2 \\ 2 \end{pmatrix},
X_{n \times p} = \begin{pmatrix} 1 & 0 \\ 0 & 1 \\ 1 & 1 \\ 1 & 0 \\ 0 & 1 \end{pmatrix}\]

### (b)
\[(X^T X) = 
\begin{pmatrix} 1 & 0 & 1 & 1 & 0 \\ 0 & 1 & 1 & 0 & 1 \end{pmatrix}
\begin{pmatrix} 1 & 0 \\ 0 & 1 \\ 1 & 1 \\ 1 & 0 \\ 0 & 1 \end{pmatrix}
= \begin{pmatrix} 3 & 1 \\ 1 & 3 \end{pmatrix}\]

\[(X^T X)^{-1} = 
\frac{1}{3 \cdot 3 - 1 \cdot 1}
\begin{pmatrix} 3 & -1 \\ -1 & 3 \end{pmatrix}
=
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix} \]

\[X^T Y =
\begin{pmatrix} 1 & 0 & 1 & 1 & 0 \\ 0 & 1 & 1 & 0 & 1 \end{pmatrix}
\begin{pmatrix} 2 \\ 1 \\ 4 \\ 2 \\ 2 \end{pmatrix}
=\begin{pmatrix} 8 \\ 7 \end{pmatrix}\]

### (c)
\[\hat{\beta} = (X^T X)^{-1}X^T Y
=\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
\begin{pmatrix} 8 \\ 7 \end{pmatrix}
= \begin{pmatrix} \frac{17}{8} \\ \frac{13}{8} \end{pmatrix}\]

### (d)
Yes it is the same.

### (e)
\[H = X(X^T X)^{-1} X^T =
\begin{pmatrix} 1 & 0 \\ 0 & 1 \\ 1 & 1 \\ 1 & 0 \\ 0 & 1 \end{pmatrix}
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
\begin{pmatrix} 1 & 0 & 1 & 1 & 0 \\ 0 & 1 & 1 & 0 & 1 \end{pmatrix} =
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} & \frac{1}{4} & \frac{3}{8} & -\frac{1}{8} \\
-\frac{1}{8} & \frac{3}{8} & \frac{1}{4} & -\frac{1}{8} & \frac{3}{8} \\
\frac{1}{4} & \frac{1}{4} & \frac{1}{2} & \frac{1}{4} & \frac{1}{4} \\
\frac{3}{8} & -\frac{1}{8} & \frac{1}{4} & \frac{3}{8} & -\frac{1}{8} \\
-\frac{1}{8} & \frac{3}{8}& \frac{1}{4}  & -\frac{1}{8} & \frac{3}{8}
\end{pmatrix}\]

### (f)
\[ \hat{\sigma ^ 2} = \frac{RSS}{n - p} \]
\[ \hat{Y} = HY =
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} & \frac{1}{4} & \frac{3}{8} & -\frac{1}{8} \\
-\frac{1}{8} & \frac{3}{8} & \frac{1}{4} & -\frac{1}{8} & \frac{3}{8} \\
\frac{1}{4} & \frac{1}{4} & \frac{1}{2} & \frac{1}{4} & \frac{1}{4} \\
\frac{3}{8} & -\frac{1}{8} & \frac{1}{4} & \frac{3}{8} & -\frac{1}{8} \\
-\frac{1}{8} & \frac{3}{8}& \frac{1}{4}  & -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
\begin{pmatrix} 2 \\ 1 \\ 4 \\ 2 \\ 2 \end{pmatrix}
= \begin{pmatrix} \frac{17}{8} \\ \frac{13}{8} \\ \frac{15}{4} \\ \frac{17}{8} \\ \frac{13}{8} \end{pmatrix}
\]
\[ RSS = \sum_{i} (\epsilon_i)^2 = 
(\epsilon_1)^2 + (\epsilon_2)^2 + (\epsilon_3)^2 + (\epsilon_4)^2 + (\epsilon_5)^2\]
\[ = (\frac{17}{8} - 2)^2 + (\frac{13}{8} - 1)^2 + (\frac{15}{4} - 4)^2 + (\frac{17}{8} - 2)^2 + (\frac{13}{8} - 2)^2 
= \frac{5}{8} \]
\[ \hat{\sigma ^ 2} = \frac{\frac{5}{8}}{5-2} = \frac{5}{24} \]
\[ \hat{\sigma} = \sqrt{\frac{5}{24}} \approx 0.456 \]

```{r, echo=F}
y_1 <- c(2, 1, 4, 2, 2)
a_1 <- c(1, 0, 1, 1, 0)
b_1 <- c(0, 1, 1, 0, 1)

mod_1 <- lm(y_1 ~ a_1 + b_1 -1)
summary(mod_1)
```

## Problem 2

### (a)
\[n = 6, 
Y_{n \times 1} = \begin{pmatrix} 2 \\ 1 \\ 4 \\ 2 \\ 2 \\ 5 \end{pmatrix},
X_{n \times p} = \begin{pmatrix} 1 & 0 \\ 0 & 1 \\ 1 & 1 \\ 1 & 0 \\ 0 & 1 \\ 1 & 1 \end{pmatrix}
\]

### (b)
\[(X^T X) = 
\begin{pmatrix} 1 & 0 & 1 & 1 & 0 & 1 \\ 0 & 1 & 1 & 0 & 1 & 1 \end{pmatrix}
\begin{pmatrix} 1 & 0 \\ 0 & 1 \\ 1 & 1 \\ 1 & 0 \\ 0 & 1 \\ 1 & 1 \end{pmatrix}
= \begin{pmatrix} 4 & 2 \\ 2 & 4 \end{pmatrix}\]

\[(X^T X)^{-1} = 
\frac{1}{4 \cdot 4 - 2 \cdot 2}
\begin{pmatrix} 4 & -2 \\ -2 & 4 \end{pmatrix}
=
\begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix} \]

\[X^T Y =
\begin{pmatrix} 1 & 0 & 1 & 1 & 0 & 1 \\ 0 & 1 & 1 & 0 & 1 & 1 \end{pmatrix}
\begin{pmatrix} 2 \\ 1 \\ 4 \\ 2 \\ 2 \\ 5 \end{pmatrix}
=\begin{pmatrix} 13 \\ 12 \end{pmatrix}\]

### (c)
\[\hat{\beta} = (X^T X)^{-1}X^T Y
=\begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix}
\begin{pmatrix} 13 \\ 12 \end{pmatrix}
= \begin{pmatrix} \frac{7}{3} \\ \frac{11}{6} \end{pmatrix}\]

### (d)
Yes it is the same.

### (e)
\[H = X(X^T X)^{-1} X^T =
\begin{pmatrix} 1 & 0 \\ 0 & 1 \\ 1 & 1 \\ 1 & 0 \\ 0 & 1 \\ 1 & 1 \end{pmatrix}
\begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix}
\begin{pmatrix} 1 & 0 & 1 & 1 & 0 & 1 \\ 0 & 1 & 1 & 0 & 1 & 1 \end{pmatrix}
=
\begin{pmatrix} 
\frac{1}{3} & -\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & -\frac{1}{6} & \frac{1}{6} \\
-\frac{1}{6} & \frac{1}{3} & \frac{1}{6} & -\frac{1}{6} & \frac{1}{3} & \frac{1}{6}    \\
\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & \frac{1}{6} & \frac{1}{6} & \frac{1}{3} \\
\frac{1}{3} & -\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & -\frac{1}{6} & \frac{1}{6} \\
-\frac{1}{6} & \frac{1}{3} & \frac{1}{6} & -\frac{1}{6} & \frac{1}{3} & \frac{1}{6}    \\
\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & \frac{1}{6} & \frac{1}{6} & \frac{1}{3} 
\end{pmatrix}
\]

### (f)
\[ \hat{\sigma ^ 2} = \frac{RSS}{n - p} \]
\[ \hat{Y} = HY =
\begin{pmatrix} 
\frac{1}{3} & -\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & -\frac{1}{6} & \frac{1}{6} \\
-\frac{1}{6} & \frac{1}{3} & \frac{1}{6} & -\frac{1}{6} & \frac{1}{3} & \frac{1}{6}    \\
\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & \frac{1}{6} & \frac{1}{6} & \frac{1}{3} \\
\frac{1}{3} & -\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & -\frac{1}{6} & \frac{1}{6} \\
-\frac{1}{6} & \frac{1}{3} & \frac{1}{6} & -\frac{1}{6} & \frac{1}{3} & \frac{1}{6}    \\
\frac{1}{6} & \frac{1}{6} & \frac{1}{3} & \frac{1}{6} & \frac{1}{6} & \frac{1}{3} 
\end{pmatrix}
\begin{pmatrix} 2 \\ 1 \\ 4 \\ 2 \\ 2 \\ 5 \end{pmatrix}\
= 
\begin{pmatrix}
\frac{7}{3} \\ \frac{11}{6} \\ \frac{25}{6} \\ \frac{7}{3} \\ \frac{11}{6} \\ \frac{25}{6}
\end{pmatrix}
\]
\[ RSS = \sum_{i} (\epsilon_i)^2 = 
(\epsilon_1)^2 + (\epsilon_2)^2 + (\epsilon_3)^2 + (\epsilon_4)^2 + (\epsilon_5)^2 + (\epsilon_6)^2\]
\[ = (\frac{7}{3} - 2)^2 + (\frac{11}{6} - 1)^2 + (\frac{25}{6} - 4)^2 + (\frac{7}{3} - 2)^2 + (\frac{11}{6} - 2)^2 + (\frac{25}{6} - 5)^2 
= \frac{5}{3} \]
\[ \hat{\sigma ^ 2} = \frac{\frac{5}{3}}{6-2} = \frac{5}{12} \]
\[ \hat{\sigma} = \sqrt{\frac{5}{12}} \approx 0.645 \]

```{r, echo=F}
y_2 <- c(2, 1, 4, 2, 2, 5)
a_2 <- c(1, 0, 1, 1, 0, 1)
b_2 <- c(0, 1, 1, 0, 1, 1)

mod_2 <- lm(y_2 ~ a_2 + b_2 -1)
summary(mod_2)
```

## Problem 3

### (a)
\[n = 7, 
Y_{n \times 1} = 
\begin{pmatrix} 
2 \\
1 \\
4 \\
2 \\
2 \\
5 \\
5 
\end{pmatrix},
X_{n \times p} = 
\begin{pmatrix}
1 & 0 \\
0 & 1 \\
1 & 1 \\
1 & 0 \\
0 & 1 \\
1 & 1 \\
1 & 1
\end{pmatrix}
\]

### (b)
\[(X^T X) =
\begin{pmatrix}
1 & 0 & 1 & 1 & 0 & 1 & 1 \\ 
0 & 1 & 1 & 0 & 1 & 1 & 1 
\end{pmatrix}
\begin{pmatrix}
1 & 0 \\
0 & 1 \\
1 & 1 \\
1 & 0 \\
0 & 1 \\
1 & 1 \\
1 & 1
\end{pmatrix}
= 
\begin{pmatrix} 
5 & 3 \\
3 & 5
\end{pmatrix}
\]

\[(X^T X)^{-1} = 
\frac{1}{5 \cdot 5 - 3 \cdot 3}
\begin{pmatrix} 
5 & -3 \\
-3 & 5
\end{pmatrix}
=
\begin{pmatrix}
\frac{5}{16} & -\frac{3}{16} \\
-\frac{3}{16} & \frac{5}{16}
\end{pmatrix}
\]

\[X^T Y =
\begin{pmatrix}
1 & 0 & 1 & 1 & 0 & 1 & 1 \\ 
0 & 1 & 1 & 0 & 1 & 1 & 1 
\end{pmatrix}
\begin{pmatrix} 
2 \\
1 \\
4 \\
2 \\
2 \\
5 \\
5 
\end{pmatrix}\
=
\begin{pmatrix}
18 \\
17
\end{pmatrix}
\]

### (c)
\[\hat{\beta} = (X^T X)^{-1}X^T Y
=
\begin{pmatrix}
\frac{5}{16} & -\frac{3}{16} \\
-\frac{3}{16} & \frac{5}{16}
\end{pmatrix}
\begin{pmatrix}
18 \\
17
\end{pmatrix}
= \begin{pmatrix}
\frac{39}{16} \\
\frac{31}{16}
\end{pmatrix}\]

### (d)
Yes it is the same.

### (e)
\[H = X(X^T X)^{-1} X^T =
\begin{pmatrix}
1 & 0 \\
0 & 1 \\
1 & 1 \\
1 & 0 \\
0 & 1 \\
1 & 1 \\
1 & 1
\end{pmatrix}
\begin{pmatrix}
\frac{5}{16} & -\frac{3}{16} \\
-\frac{3}{16} & \frac{5}{16}
\end{pmatrix}
\begin{pmatrix}
1 & 0 & 1 & 1 & 0 & 1 & 1 \\ 
0 & 1 & 1 & 0 & 1 & 1 & 1 
\end{pmatrix}
=
\begin{pmatrix} 
\frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{1}{8} \\
-\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & -\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & \frac{1}{8} \\
\frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{4} \\
\frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{1}{8} \\
-\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & -\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & \frac{1}{8} \\
\frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{4} \\
\frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{4} \\
\end{pmatrix}
\]

### (f)
\[ \hat{\sigma ^ 2} = \frac{RSS}{n - p} \]
\[ \hat{Y} = HY =
\begin{pmatrix} 
\frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{1}{8} \\
-\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & -\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & \frac{1}{8} \\
\frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{4} \\
\frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{5}{16} & -\frac{3}{16} & \frac{1}{8} & \frac{1}{8} \\
-\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & -\frac{3}{16} & \frac{5}{16} & \frac{1}{8} & \frac{1}{8} \\
\frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{4} \\
\frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{8} & \frac{1}{8} & \frac{1}{4} & \frac{1}{4} \\
\end{pmatrix}
\begin{pmatrix} 
2 \\
1 \\
4 \\
2 \\
2 \\
5 \\
5 
\end{pmatrix}
= 
\begin{pmatrix}
\frac{39}{16} \\ \frac{31}{16} \\ \frac{35}{8} \\ \frac{39}{16} \\ \frac{31}{16} \\ \frac{35}{8} \\ \frac{35}{8}
\end{pmatrix}
\]
\[ RSS = \sum_{i} (\epsilon_i)^2 = 
(\epsilon_1)^2 + (\epsilon_2)^2 + (\epsilon_3)^2 + (\epsilon_4)^2 + (\epsilon_5)^2 + (\epsilon_6)^2 + (\epsilon_7)^2\]
\[ = (\frac{39}{16} - 2)^2 + (\frac{31}{16} - 1)^2 + (\frac{35}{8} - 4)^2 + (\frac{39}{16} - 2)^2 + (\frac{31}{16} - 2)^2 + (\frac{35}{8} - 5)^2 + (\frac{35}{8} - 5)^2
= \frac{35}{16} \]
\[ \hat{\sigma ^ 2} = \frac{\frac{35}{16}}{7-2} = \frac{7}{16} \]
\[ \hat{\sigma} = \sqrt{\frac{7}{16}} \approx 0.661 \]

```{r, echo=F}
y_3 <- c(2, 1, 4, 2, 2, 5, 5)
a_3 <- c(1, 0, 1, 1, 0, 1, 1)
b_3 <- c(0, 1, 1, 0, 1, 1, 1)

mod_3 <- lm(y_3 ~ a_3 + b_3 -1)
summary(mod_3)
```

## Problem 4

### (a)
\[n = 3, 
Y_{n \times 1} = 
\begin{pmatrix} 
2 \\
1.5 \\
4.5 
\end{pmatrix},
X_{n \times p} = 
\begin{pmatrix}
1 & 0 \\
0 & 1 \\
1 & 1 \\
\end{pmatrix}
\]

### (b)
\[(X^T X) =
\begin{pmatrix}
1 & 0 & 1 \\ 
0 & 1 & 1 
\end{pmatrix}
\begin{pmatrix}
1 & 0 \\
0 & 1 \\
1 & 1 \\
\end{pmatrix}
= 
\begin{pmatrix} 
2 & 1 \\
1 & 2
\end{pmatrix}
\]

\[(X^T X)^{-1} = 
\frac{1}{2 \cdot 2 - 1 \cdot 1}
\begin{pmatrix} 
2 & -1 \\
-1 & 2
\end{pmatrix}
=
\begin{pmatrix}
\frac{2}{3} & -\frac{1}{3} \\
-\frac{1}{3} & \frac{2}{3}
\end{pmatrix}
\]

\[X^T Y =
\begin{pmatrix}
1 & 0 & 1 \\ 
0 & 1 & 1 
\end{pmatrix}
\begin{pmatrix} 
2 \\
1.5 \\
4.5 
\end{pmatrix}
=
\begin{pmatrix}
6.5 \\
6
\end{pmatrix}
\]

### (c)
\[\hat{\beta} = (X^T X)^{-1}X^T Y
=
\begin{pmatrix}
\frac{2}{3} & -\frac{1}{3} \\
-\frac{1}{3} & \frac{2}{3}
\end{pmatrix}
\begin{pmatrix}
6.5 \\
6
\end{pmatrix}
= \begin{pmatrix}
\frac{7}{3} \\
\frac{11}{6}
\end{pmatrix}\]

### (d)
Yes it is the same.

### (e)
\[H = X(X^T X)^{-1} X^T =
\begin{pmatrix}
1 & 0 \\
0 & 1 \\
1 & 1 \\
\end{pmatrix}
\begin{pmatrix}
\frac{2}{3} & -\frac{1}{3} \\
-\frac{1}{3} & \frac{2}{3}
\end{pmatrix}
\begin{pmatrix}
1 & 0 & 1 \\ 
0 & 1 & 1 
\end{pmatrix}
=
\begin{pmatrix} 
\frac{2}{3} & -\frac{1}{3} & \frac{1}{3} \\
-\frac{1}{3} & \frac{2}{3} & \frac{1}{3} \\
\frac{1}{3} & \frac{1}{3} & \frac{2}{3}
\end{pmatrix}
\]

### (f)
\[ \hat{\sigma ^ 2} = \frac{RSS}{n - p} \]
\[ \hat{Y} = HY =
\begin{pmatrix} 
\frac{2}{3} & -\frac{1}{3} & \frac{1}{3} \\
-\frac{1}{3} & \frac{2}{3} & \frac{1}{3} \\
\frac{1}{3} & \frac{1}{3} & \frac{2}{3}
\end{pmatrix}
\begin{pmatrix} 
2 \\
1.5 \\
4.5 
\end{pmatrix}
= 
\begin{pmatrix}
\frac{7}{3} \\ \frac{11}{6} \\ \frac{25}{6}
\end{pmatrix}
\]
\[ RSS = \sum_{i} (\epsilon_i)^2 = 
(\epsilon_1)^2 + (\epsilon_2)^2 + (\epsilon_3)^2 \]
\[ = (\frac{7}{3} - 2)^2 + (\frac{11}{6} - 1.5)^2 + (\frac{25}{6} - 4.5)^2 
= \frac{1}{3} \]
\[ \hat{\sigma ^ 2} = \frac{\frac{1}{3}}{3-2} = \frac{1}{3} \]
\[ \hat{\sigma} = \sqrt{\frac{1}{3}} \approx 0.577 \]

```{r, echo=F}
y_4 <- c(2, 1.5, 4.5)
a_4 <- c(1, 0, 1)
b_4 <- c(0, 1, 1)

mod_4 <- lm(y_4 ~ a_4 + b_4 -1)
summary(mod_4)
```

### (g)
The hat matrix is always symmetrical.