---
title: "HW3"
author: "Andrew Shao"
date: "2024-09-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Problem 1


### (a)
\[ CI = \hat{A} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.2, n = 5, p = 2 \] 
  \[ x_{new} = \begin{pmatrix} 1 \\ 0 \end{pmatrix} \]
  \[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
  \begin{pmatrix} 1 & 0 \end{pmatrix}
  \begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
  \begin{pmatrix} 1 \\ 0 \end{pmatrix}
  = \frac{3}{8} \]
\[ \hat{A} = \frac{17}{8}, t_{\frac{0.2}{2}, 5-2} = t_{0.1, 3} = 1.638, \hat{\sigma} = \sqrt{\frac{5}{24}} \]
\[ CI = \frac{17}{8} \pm 1.638 \sqrt{\frac{5}{24}} \sqrt{\frac{3}{8}} = [1.667, 2.583] \]


### (b)
\[ CI = \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.2, n = 5, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 0 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 0 & 1 \end{pmatrix}
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
\begin{pmatrix} 0 \\ 1 \end{pmatrix}
= \frac{3}{8} \]
\[ \hat{B} = \frac{13}{8}, t_{\frac{0.2}{2}, 5-2} = t_{0.1, 3} = 1.638, \hat{\sigma} = \sqrt{\frac{5}{24}} \]
\[ CI = \frac{13}{8} \pm 1.638 \sqrt{\frac{5}{24}} \sqrt{\frac{3}{8}} = [1.167, 2.083] \]


### (c)
\[ CI = \hat{A} + \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.2, n = 5, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & 1 \end{pmatrix}
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
\begin{pmatrix} 1 \\ 1 \end{pmatrix}
= \frac{1}{2} \]
\[ \hat{A} = \frac{17}{8}, \hat{B} = \frac{13}{8}, t_{\frac{0.2}{2}, 5-2} = t_{0.1, 3} = 1.638, \hat{\sigma} = \sqrt{\frac{5}{24}} \]
\[ CI = \frac{17}{8} + \frac{13}{8} \pm 1.638 \sqrt{\frac{5}{24}} \sqrt{\frac{1}{2}} = [3.221, 4.279] \]


### (d)
\[ CI = \hat{A} - \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.2, n = 5, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ -1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & -1 \end{pmatrix}
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
\begin{pmatrix} 1 \\ -1 \end{pmatrix}
= 1 \]
\[ \hat{A} = \frac{17}{8}, \hat{B} = \frac{13}{8}, t_{\frac{0.2}{2}, 5-2} = t_{0.1, 3} = 1.638, \hat{\sigma} = \sqrt{\frac{5}{24}} \]
\[ CI = \frac{17}{8} - \frac{13}{8} \pm 1.638 \sqrt{\frac{5}{24}} \sqrt{1} = [-0.248, 1.248] \]


### (e)
\[ CI = \hat{A} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{1 + x_{new}^T (X^T X)^{-1}x_{new}} \]
\[ \alpha = 0.2, n = 5, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ 0 \end{pmatrix} \]
\[ x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & 0 \end{pmatrix}
\begin{pmatrix} \frac{3}{8} & -\frac{1}{8} \\ -\frac{1}{8} & \frac{3}{8} \end{pmatrix}
\begin{pmatrix} 1 \\ 0 \end{pmatrix}
= \frac{3}{8} \]
\[ \hat{A} = \frac{17}{8}, t_{\frac{0.2}{2}, 5-2} = t_{0.1, 3} = 1.638, \hat{\sigma} = \sqrt{\frac{5}{24}} \]
\[ CI = \frac{17}{8} \pm 1.638 \sqrt{\frac{5}{24}} \sqrt{1 + \frac{3}{8}} = [1.248, 3.002] \]

### (f)
```{r}
y_1 <- c(2, 1, 4, 2, 2)
a_1 <- c(1, 0, 1, 1, 0)
b_1 <- c(0, 1, 1, 0, 1)

mod_1 <- lm(y_1 ~ a_1 + b_1 -1)
predict(mod_1, data.frame(a_1 = 1, b_1 = 0), interval = 'confidence', level = 0.8)
```
```{r}
predict(mod_1, data.frame(a_1 = 0, b_1 = 1), interval = 'confidence', level = 0.8)
```
```{r}
predict(mod_1, data.frame(a_1 = 1, b_1 = 1), interval = 'confidence', level = 0.8)
```
```{r}
predict(mod_1, data.frame(a_1 = 1, b_1 = -1), interval = 'confidence', level = 0.8)
```
```{r}
predict(mod_1, data.frame(a_1 = 1, b_1 = 0), interval = 'prediction', level = 0.8)
```


## Problem 2


### (a)
\[ CI = \hat{A} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.1, n = 6, p = 2 \] 
  \[ x_{new} = \begin{pmatrix} 1 \\ 0 \end{pmatrix} \]
  \[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
  \begin{pmatrix} 1 & 0 \end{pmatrix}
  \begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix}
  \begin{pmatrix} 1 \\ 0 \end{pmatrix}
  = \frac{1}{3} \]
\[ \hat{A} = \frac{7}{3}, t_{\frac{0.1}{2}, 6-2} = t_{0.05, 4} = 2.132, \hat{\sigma} = \sqrt{\frac{5}{12}} \]
\[ CI = \frac{7}{3} \pm 2.132 \sqrt{\frac{5}{12}} \sqrt{\frac{1}{3}} = [1.539, 3.128] \]


### (b)
\[ CI = \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.1, n = 6, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 0 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 0 & 1 \end{pmatrix}
\begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix}
\begin{pmatrix} 0 \\ 1 \end{pmatrix}
= \frac{1}{3} \]
\[ \hat{B} = \frac{11}{6}, t_{\frac{0.1}{2}, 6-2} = t_{0.05, 4} = 2.132, \hat{\sigma} = \sqrt{\frac{5}{12}} \]
\[ CI = \frac{11}{6} \pm 2.132 \sqrt{\frac{5}{12}} \sqrt{\frac{1}{3}} = [1.039, 2.628] \]


### (c)
\[ CI = \hat{A} + \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.1, n = 6, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & 1 \end{pmatrix}
\begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix}
\begin{pmatrix} 1 \\ 1 \end{pmatrix}
= \frac{1}{3} \]
\[ \hat{A} = \frac{7}{3}, \hat{B} = \frac{11}{6}, t_{\frac{0.1}{2}, 6-2} = t_{0.05, 4} = 2.132, \hat{\sigma} = \sqrt{\frac{5}{12}} \]
\[ CI = \frac{7}{3} + \frac{11}{6} \pm 2.132 \sqrt{\frac{5}{12}} \sqrt{\frac{1}{3}} = [3.372, 4.961] \]


### (d)
\[ CI = \hat{A} - \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.1, n = 6, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ -1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & -1 \end{pmatrix}
\begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix}
\begin{pmatrix} 1 \\ -1 \end{pmatrix}
= 1 \]
\[ \hat{A} = \frac{7}{3}, \hat{B} = \frac{11}{6}, t_{\frac{0.1}{2}, 6-2} = t_{0.05, 4} = 2.132, \hat{\sigma} = \sqrt{\frac{5}{12}} \]
\[ CI = \frac{7}{3} - \frac{11}{6} \pm 2.132 \sqrt{\frac{5}{12}} \sqrt{1} = [-0.876, 1.876] \]


### (e)
\[ CI = \hat{A} + \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{1 + x_{new}^T (X^T X)^{-1}x_{new}} \]
\[ \alpha = 0.1, n = 6, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ 1 \end{pmatrix} \]
\[ x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & 1 \end{pmatrix}
\begin{pmatrix} \frac{1}{3} & -\frac{1}{6} \\ -\frac{1}{6} & \frac{1}{3} \end{pmatrix}
\begin{pmatrix} 1 \\ 1 \end{pmatrix}
= \frac{1}{3} \]
\[ \hat{A} = \frac{7}{3}, \hat{B} = \frac{11}{6}, t_{\frac{0.1}{2}, 6-2} = t_{0.05, 4} = 2.132, \hat{\sigma} = \sqrt{\frac{5}{12}} \]
\[ CI = \frac{7}{3} + \frac{11}{6} \pm 2.132 \sqrt{\frac{5}{12}} \sqrt{1 + \frac{1}{3}} = [2.578, 5.756] \]

### (f)
```{r}
y_2 <- c(2, 1, 4, 2, 2, 5)
a_2 <- c(1, 0, 1, 1, 0, 1)
b_2 <- c(0, 1, 1, 0, 1, 1)

mod_2 <- lm(y_2 ~ a_2 + b_2 -1)
predict(mod_2, data.frame(a_2 = 1, b_2 = 0), interval = 'confidence', level = 0.9)
```
```{r}
predict(mod_2, data.frame(a_2 = 0, b_2 = 1), interval = 'confidence', level = 0.9)
```
```{r}
predict(mod_2, data.frame(a_2 = 1, b_2 = 1), interval = 'confidence', level = 0.9)
```
```{r}
predict(mod_2, data.frame(a_2 = 1, b_2 = -1), interval = 'confidence', level = 0.9)
```
```{r}
predict(mod_2, data.frame(a_2 = 1, b_2 = 1), interval = 'prediction', level = 0.9)
```


## Problem 3


### (a)
\[ CI = \hat{A} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.05, n = 7, p = 2 \] 
  \[ x_{new} = \begin{pmatrix} 1 \\ 0 \end{pmatrix} \]
  \[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
  \begin{pmatrix} 1 & 0 \end{pmatrix}
  \begin{pmatrix} \frac{5}{16} & -\frac{3}{16} \\ -\frac{3}{16} & \frac{5}{16} \end{pmatrix}
  \begin{pmatrix} 1 \\ 0 \end{pmatrix}
  = \frac{5}{16} \]
\[ \hat{A} = \frac{39}{16}, t_{\frac{0.05}{2}, 7-2} = t_{0.025, 5} = 2.571, \hat{\sigma} = \sqrt{\frac{7}{16}} \]
\[ CI = \frac{39}{16} \pm 2.571 \sqrt{\frac{7}{16}} \sqrt{\frac{5}{16}} = [1.487, 3.388] \]


### (b)
\[ CI = \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.05, n = 7, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 0 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 0 & 1 \end{pmatrix}
\begin{pmatrix} \frac{5}{16} & -\frac{3}{16} \\ -\frac{3}{16} & \frac{5}{16} \end{pmatrix}
\begin{pmatrix} 0 \\ 1 \end{pmatrix}
= \frac{5}{16} \]
\[ \hat{B} = \frac{31}{16}, t_{\frac{0.05}{2}, 7-2} = t_{0.025, 5} = 2.571, \hat{\sigma} = \sqrt{\frac{7}{16}} \]
\[ CI = \frac{31}{16} \pm 2.571 \sqrt{\frac{7}{16}} \sqrt{\frac{5}{16}} = [0.987, 2.888] \]


### (c)
\[ CI = \hat{A} + \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.05, n = 7, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & 1 \end{pmatrix}
\begin{pmatrix} \frac{5}{16} & -\frac{3}{16} \\ -\frac{3}{16} & \frac{5}{16} \end{pmatrix}
\begin{pmatrix} 1 \\ 1 \end{pmatrix}
= \frac{1}{4} \]
\[ \hat{A} = \frac{39}{16}, \hat{B} = \frac{31}{16}, t_{\frac{0.05}{2}, 7-2} = t_{0.025, 5} = 2.571, \hat{\sigma} = \sqrt{\frac{7}{16}} \]
\[ CI = \frac{39}{16} + \frac{31}{16} \pm 2.571 \sqrt{\frac{7}{16}} \sqrt{\frac{1}{4}} = [3.525, 5.225] \]


### (d)
\[ CI = \hat{A} - 2 \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.05, n = 7, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ -2 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & -2 \end{pmatrix}
\begin{pmatrix} \frac{5}{16} & -\frac{3}{16} \\ -\frac{3}{16} & \frac{5}{16} \end{pmatrix}
\begin{pmatrix} 1 \\ -2 \end{pmatrix}
= \frac{37}{16} \]
\[ \hat{A} = \frac{39}{16}, \hat{B} = \frac{31}{16}, t_{\frac{0.05}{2}, 7-2} = t_{0.025, 5} = 2.571, \hat{\sigma} = \sqrt{\frac{7}{16}} \]
\[ CI = \frac{39}{16} - 2 \cdot \frac{31}{16} \pm 2.571 \sqrt{\frac{7}{16}} \sqrt{\frac{37}{16}} = [-4.023, 1.148] \]


### (e)
\[ CI = \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{1 + x_{new}^T (X^T X)^{-1}x_{new}} \]
\[ \alpha = 0.05, n = 7, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 0 \\ 1 \end{pmatrix} \]
\[ x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 0 & 1 \end{pmatrix}
\begin{pmatrix} \frac{5}{16} & -\frac{3}{16} \\ -\frac{3}{16} & \frac{5}{16} \end{pmatrix}
\begin{pmatrix} 0 \\ 1 \end{pmatrix}
= \frac{5}{16} \]
\[ \hat{B} = \frac{31}{16}, t_{\frac{0.05}{2}, 7-2} = t_{0.025, 5} = 2.571, \hat{\sigma} = \sqrt{\frac{7}{16}} \]
\[ CI = \frac{31}{16} \pm 2.571 \sqrt{\frac{7}{16}} \sqrt{1 + \frac{5}{16}} = [-0.010, 3.885] \]

### (f)
```{r}
y_3 <- c(2, 1, 4, 2, 2, 5, 5)
a_3 <- c(1, 0, 1, 1, 0, 1, 1)
b_3 <- c(0, 1, 1, 0, 1, 1, 1)

mod_3 <- lm(y_3 ~ a_3 + b_3 -1)
predict(mod_3, data.frame(a_3 = 1, b_3 = 0), interval = 'confidence', level = 0.95)
```
```{r}
predict(mod_3, data.frame(a_3 = 0, b_3 = 1), interval = 'confidence', level = 0.95)
```
```{r}
predict(mod_3, data.frame(a_3 = 1, b_3 = 1), interval = 'confidence', level = 0.95)
```
```{r}
predict(mod_3, data.frame(a_3 = 1, b_3 = -2), interval = 'confidence', level = 0.95)
```
```{r}
predict(mod_3, data.frame(a_3 = 0, b_3 = 1), interval = 'prediction', level = 0.95)
```


## Problem 4


### (a)
\[ CI = \hat{A} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.3, n = 3, p = 2 \] 
  \[ x_{new} = \begin{pmatrix} 1 \\ 0 \end{pmatrix} \]
  \[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
  \begin{pmatrix} 1 & 0 \end{pmatrix}
  \begin{pmatrix} \frac{2}{3} & -\frac{1}{3} \\ -\frac{1}{3} & \frac{2}{3} \end{pmatrix}
  \begin{pmatrix} 1 \\ 0 \end{pmatrix}
  = \frac{2}{3} \]
\[ \hat{A} = \frac{7}{3}, t_{\frac{0.3}{2}, 3-2} = t_{0.15, 1} = 1.963, \hat{\sigma} = \sqrt{\frac{1}{3}} \]
\[ CI = \frac{7}{3} \pm 1.963 \sqrt{\frac{1}{3}} \sqrt{\frac{2}{3}} = [1.408, 3.259] \]


### (b)
\[ CI = \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.3, n = 3, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 0 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 0 & 1 \end{pmatrix}
\begin{pmatrix} \frac{2}{3} & -\frac{1}{3} \\ -\frac{1}{3} & \frac{2}{3} \end{pmatrix}
\begin{pmatrix} 0 \\ 1 \end{pmatrix}
= \frac{2}{3} \]
\[ \hat{B} = \frac{11}{6}, t_{\frac{0.3}{2}, 3-2} = t_{0.15, 1} = 1.963, \hat{\sigma} = \sqrt{\frac{1}{3}} \]
\[ CI = \frac{11}{6} \pm 1.963 \sqrt{\frac{1}{3}} \sqrt{\frac{2}{3}} = [0.908, 2.759] \]


### (c)
\[ CI = \hat{A} + \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.3, n = 3, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ 1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & 1 \end{pmatrix}
\begin{pmatrix} \frac{2}{3} & -\frac{1}{3} \\ -\frac{1}{3} & \frac{2}{3} \end{pmatrix}
\begin{pmatrix} 1 \\ 1 \end{pmatrix}
= \frac{2}{3} \]
\[ \hat{A} = \frac{7}{3}, \hat{B} = \frac{11}{6}, t_{\frac{0.3}{2}, 3-2} = t_{0.15, 1} = 1.963, \hat{\sigma} = \sqrt{\frac{1}{3}} \]
\[ CI = \frac{7}{3} + \frac{11}{6} \pm 1.963 \sqrt{\frac{1}{3}} \sqrt{\frac{2}{3}} = [3.241, 5.092] \]


### (d)
\[ CI = \hat{A} - \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{d_1} \]
\[ \alpha = 0.3, n = 3, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ -1 \end{pmatrix} \]
\[ d_1 = x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & -1 \end{pmatrix}
\begin{pmatrix} \frac{2}{3} & -\frac{1}{3} \\ -\frac{1}{3} & \frac{2}{3} \end{pmatrix}
\begin{pmatrix} 1 \\ -1 \end{pmatrix}
= 2 \]
\[ \hat{A} = \frac{7}{3}, \hat{B} = \frac{11}{6}, t_{\frac{0.3}{2}, 3-2} = t_{0.15, 1} = 1.963, \hat{\sigma} = \sqrt{\frac{1}{3}} \]
\[ CI = \frac{7}{3} - \frac{11}{6} \pm 1.963 \sqrt{\frac{1}{3}} \sqrt{2} = [-1.102, 2.102] \]


### (e)
\[ CI = \hat{A} + \hat{B} \pm t_{\frac{\alpha}{2}, n-p} \hat{\sigma} \sqrt{1 + x_{new}^T (X^T X)^{-1}x_{new}} \]
\[ \alpha = 0.3, n = 3, p = 2 \] 
\[ x_{new} = \begin{pmatrix} 1 \\ 1 \end{pmatrix} \]
\[ x_{new}^T (X^T X)^{-1}x_{new}  =
\begin{pmatrix} 1 & 1 \end{pmatrix}
\begin{pmatrix} \frac{2}{3} & -\frac{1}{3} \\ -\frac{1}{3} & \frac{2}{3} \end{pmatrix}
\begin{pmatrix} 1 \\ 1 \end{pmatrix}
= \frac{2}{3} \]
\[ \hat{A} = \frac{7}{3}, \hat{B} = \frac{11}{6}, t_{\frac{0.3}{2}, 3-2} = t_{0.15, 1} = 1.963, \hat{\sigma} = \sqrt{\frac{1}{3}} \]
\[ CI = \frac{7}{3} + \frac{11}{6} \pm 1.963 \sqrt{\frac{1}{3}} \sqrt{1 + \frac{2}{3}} = [2.704, 5.630] \]

### (f)
With more measurements the confidence intervals are smaller. With less balls being measured the confidence interval is also smaller. In general, the confidence intervals are quite large so more measurements are needed for precise point estimations using this method.